Classless Java: Tuning Java Interfaces

### 2015-11-29 paper/sections/Appendix.tex
Proof.

\subsection{Proof of LEMMA 1}\label{subsec:proof1}
\textbf{LEMMA 1. }
Assume that
\begin{align*}
\II_0\ &=\ \mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC\\
[\![\II_0]\!]\ &=\ \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \methods' \cC
\end{align*}
If $\II_0$ satisfies $\dom(\C_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, then $[\![\II_0]\!]$ satisfies $\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$.
\begin{proof}
It suffices to prove $\dom([\![\II_0]\!])=\dom(\C_0)\cup\dom(\methods')$.
\begin{itemize}
\item If $m\notin\dom(\methods')$, by the definition of $\mBody$, it is obtained from $\override$. The first argument of $\override$, namely $\methods\cup\methods'(m)$, is equal to $\methods(m)$, and the second argument is not changed as well, thus the result of $\override$ is not changed during translation. Hence $m\in\dom([\![\II_0]\!])$ iff $m\in\dom(\C_0)$.
\item If $m\in\dom(\methods')$, we are to prove $m\in\dom([\![\II_0]\!])$. So if $m$ is the $\QM{of}$ method generated by $\ofMethod$, its $\mBody$ value is well defined, by the rule $\override(\method,\none)=\method$. On the other hand, if $m$ is generated by $\otherMethod$, namely $m$ is a $\QM{with-}$ or setter method, we can also get its $\mBody$ value, since in the definition of $\otherMethod$, we can see $\isWith$ and $\isSetter$ ensure the compatible subtyping relationship. Hence $m\in\dom([\![\II_0]\!])$ as well.
\end{itemize}
\end{proof}

\subsection{Proof of LEMMA 2}\label{subsec:proof2}
\textbf{LEMMA 2. }
If $\II_0$ OK, then $[\![\II_0]\!]$ OK.
\begin{proof}
By the rule \rn{t-Intf} in Figure~\ref{ET}, we divide it into two parts.

@@

Since $\II_0$ OK, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(\C_0)$.

\haoyuan{How to more explicitly say the default methods are ok in the new interface table, so that it is sufficient to only prove the generated ofMethod is ok?}

@@

\haoyuan{The order of conditions in \rn{t-Obj} is not clear below.}

\begin{itemize}
\item For field typing,
\[\Gamma(\QM_ m_i)=\C_i\subtype \C_i\]
\item For method typing of getters,
\[\Gamma, m_i:\C_i, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash m_i\in \C_i=\C^{\mh_i}\]
\item For method typing of \QM{with-} methods, by \rn{t-StaticInvk}, \haoyuan{how to do typing for the generated method itself?}
\[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0, \es_i:\Cs_i, \Gamma^{\mh_i}\vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0=C^{\mh_i}\]
\item For method typing of setter methods, \haoyuan{what about the \QM{with} method?}
\[\Gamma, \QM{this}:\C_0, \Gamma^{\mh_i}\vdash\QM{this}\in \C_0=\C^{\mh_i}\]
\item To prove that $\forall i$, $\mh_i\subtype\mBody(m^{\mh_i}, \C_0)$,
 \begin{itemize}
 \item For getters, \[\C_i\ m_i\oR\cR\QM; \in \fieldsFunc(\C_0) \Rightarrow \C_i\ m_i\oR\cR\QM; \in \dom(\C_0)\]
 \item For \QM{with-} methods,
  \[\begin{array}{l}
  \mBody(\QM{with#}\m_i,\C_0) \mbox{ is of form }\mh\QM; \\ \valid(\C_0)
  \end{array}\Rightarrow \isWith\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
 \item For setters,
  \[\begin{array}{l}
  \mBody(\m_i,\C_0) \mbox{ is of form }\mh\QM; \\ \valid(\C_0)
  \end{array}\Rightarrow \isSetter\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
 \end{itemize}
\item To prove that such a created object indeed implements all the abstract methods in $\dom(\C_0)$, we simply refer to $\valid(\C_0)$, since it guarantees each abstract method $\method$ to satisfy $\isField$, $\isWith$ or $\isSetter$. But that object includes all implementations for those cases, hence it is of type $\C_0$ by \rn{t-Obj}.
\end{itemize}

\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\methods')$$
And actually it is guaranteed by \textbf{LEMMA 1}.

\end{proof}

\subsection{Proof of THEOREM}\label{subsec:proof3}
\textbf{THEOREM. }
For a given $\II_0\ldots\II_n$ interface table such that
$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
$[\![\II_0]\!]\II_1\ldots\II_n$
$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
\begin{proof}
\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, if some $\II$ is not a subtype of $\II_0$, the generated code in translation has no way to affect the domain of $\II$, which finishes our proof.
\end{proof}

>>>>

\textbf{THEOREM. }
For a given $\II_0\ldots\II_n$ interface table such that
$\forall\II\in\II_0\ldots\II_n, \II$ OK,\\*
if $\II_0$ has \textbf{@ObjWeak},
$\valid(\C_0)$  and $\QM{of}\notin\dom(\C_0)$,
then in the interface table
$[\![\II_0]\!]\II_1\ldots\II_n$\\*
$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ $\II$ OK.\\

\noindent For convenience, assume that
\begin{align*}
\II_0\ &=\ \weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC\\
[\![\II_0]\!]\ &=\ \emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \method' \cC
\end{align*}

\subsection{LEMMA 1 and Proof}\label{subsec:lemma1}

\textbf{LEMMA 1. }
For any expression $e$ under an interface table $\II_0\ldots\II_n$ with $\Gamma\vdash e\in\C$, if $\II_0$ has \textbf{@ObjWeak} and $[\![\II_0]\!]$ is well-defined, then under the interface table $[\![\II_0]\!]\II_1\ldots\II_n$, $\Gamma\vdash e\in\_<:\C$.
\begin{proof}
The proof is based on induction. By the grammar shown in Figure~\ref{Grammar}, there are three cases for an arbitrary expression $e$:
\begin{itemize}
\item A variable or a field update. The type preservation for $e$ is ensured by induction.
\item A method call (normal, static or super). Such a method won't be ``removed'' by the translation and is still there, and the types remain unchanged. The only work \textbf{@ObjWeak} does is adding a static method $\QM{of}$ to the interface, however, a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so adding $\QM{of}$ method has no way to affect any method call.
\item An object creation. Adding the $\QM{of}$ method doesn't introduce unimplemented methods to an interface, moreover, the static method is not inheritable, hence after translation such an object creation still type checks and has the right type by induction.
\end{itemize}
\end{proof}

\subsection{LEMMA 2 and Proof}\label{subsec:lemma2}

\textbf{LEMMA 2. }
If $\II_0$ has \textbf{@ObjWeak} and is OK, then $[\![\II_0]\!]$ OK.
\begin{proof}

By the rule \rn{t-Intf} in Figure~\ref{ET}, we divide the proof into two parts.

@@

Since $\II_0$ OK, and by \textbf{LEMMA 1}, all the existing default and static methods are well typed in $[\![\II_0]\!]$, except for the new method \QM{of}. It suffices to prove that it still holds for $\ofMethod(\C_0)$.

@@

\begin{itemize}
\item All field initializations are type correct. By the definition of $\ofMethod(\C_0)$ in Section~\ref{subsec:ofmethod}, the fields $m_1,\ldots,m_n$ are initialized by $\QM{of}$'s arguments, and types are compatible.
\item All method bodies are well-typed.
    \begin{itemize}
    \item Typing of the $i$-getter $m_i$. \[\Gamma, m_i:\C_i, \QM{this}:\C_0 \vdash m_i\in \C_i\]
        We know that $\C_i=\C^{\mh_i}$ since the $i$-th getter has its return type the same as the corresponding field $m_i$.
    \item Typing of the \QM{with-} method of an arbitrary field $m_i$. By Section~\ref{subsec:ofmethod}, if the \QM{with-} method of $m_i$ is well-defined, it has the form \[\C_0\ \QM{with#}\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{}\QM{return}\ \C_0\QM{.of(}\es_i\QM{);\}}\]
        $\es_i$ is obtained by replacing $m_i$ with $\QM{_val}$ in the list of fields, and since they have the same type $\C_i$, the arguments $\es_i$ are compatible with $\C_0.\QM{of}$ method. Hence \[\Gamma, \overline{m_i:\C_i}, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash \C_0\QM{.of}\oR\es_i\cR\in \C_0\]
        We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM{with#}\m_i$ shown as above.
    \item Typing of the $i$-setter \QM_$m_i$. If the \QM_$m_i$ method is well-defined, it has the form
        \[\C_0\ \QM_\m_i\oR \C_i\ \QM{_val}\cR\ \QM{\{} \m_i\QM{= _val;return this;\}}\]
        By \rn{t-Update}, the assignment ``$\m_i\QM{= _val;}$'' is correct since $\m_i$ and $\QM{_val}$ have the same type $I_i$, and the return type is decided by \QM{this}. \[\Gamma, \QM{this}:\C_0,\ \QM{_val} : \C_i \vdash\QM{this}\in \C_0\]
        We know that $\C_0=\C^{\mh_i}$ by the return type of $\QM_\m_i$ shown as above.\\
        \haoyuan{Can we write $\Gamma, \QM{this}:\C_0,\ \QM{_val} : \C_i, \C^{\mh_i} = \C_0 \vdash\QM{this}\in \C_0 = \C^{\mh_i}$?}\\
        \haoyuan{Can we write $\QM{\{} \m_i\QM{= _val;return this;\}} \in \C_0$?}
    \end{itemize}
\item All method headers are valid with respect to the domain of $\C_0$. Namely $$\sigvalid(\mh_1\ldots\mh_n,I)$$
    \haoyuan{TODO}
\item All abstract methods in the domain of $\C_0$ have been implemented. Namely $$\alldefined(\mh_1\ldots\mh_n,I)$$
    Here we simply refer to $\valid(\C_0)$, since it guarantees each abstract method to satisfy $\isField$, $\isWith$ or $\isSetter$. But that object includes all implementations for those cases, hence it is of type $\C_0$ by \rn{t-Obj}.
\end{itemize}

\noindent\textbf{Part II.} Next we check that in $[\![\II_0]\!]$, $$\dom([\![\II_0]\!])=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)\cup\dom(\method')$$
Since $\II_0$ OK, we have $\dom(\II_0)=\dom(\C_1)\cup\ldots\cup\dom(\C_n)\cup\dom(\methods)$, and hence it is equivalent to prove $$\dom([\![\II_0]\!])=\dom(\C_0)\cup\dom(\method')$$
This is obvious since a pre-condition of the translation is $\QM{of}\notin\dom(\C_0)$, so $\method'$ doesn't overlap with $\dom(\C_0)$. The definition of $\dom$ is based on $\mBody$, and here the new domain $\dom([\![\II_0]\!])$ is only an extension to $\dom(\C_0)$ with the \QM{of} method, namely $\method'$, which finishes our proof.
\end{proof}

\subsection{Notes}

\begin{itemize}
\item To prove that $\forall i$, $\mh_i\subtype\mBody(m^{\mh_i}, \C_0)$,
 \begin{itemize}
 \item For getters, \[\C_i\ m_i\oR\cR\QM; \in \fieldsFunc(\C_0) \Rightarrow \C_i\ m_i\oR\cR\QM; \in \dom(\C_0)\]
 \item For \QM{with-} methods,\marco{ use if-then or other english forms instead of the logic arrow}
  \[\begin{array}{l}
  \mBody(\QM{with#}\m_i,\C_0) \mbox{ is of form }\mh\QM; \\ \valid(\C_0)
  \end{array}\Rightarrow \isWith\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
 \item For setters,
  \[\begin{array}{l}
  \mBody(\m_i,\C_0) \mbox{ is of form }\mh\QM; \\ \valid(\C_0)
  \end{array}\Rightarrow \isSetter\Rightarrow \C_0\subtype\mBody(m^{\mh_i}, \C_0)\]
 \end{itemize}
\end{itemize}

\subsection{Proof of THEOREM}\label{subsec:proof3}
\textbf{THEOREM. }
For a given $\II_0\ldots\II_n$ interface table such that
$\forall\II\in\II_0\ldots\II_n, \II$ OK, then in the interface table
$[\![\II_0]\!]\II_1\ldots\II_n$
$\forall\II\in[\![\II_0]\!]\II_1\ldots\II_n$ either $\II$ OK or $\II$ is a subtype of $\II_0$.
\begin{proof}
\textbf{LEMMA 2} already proves that $[\![\II_0]\!]$ is OK. On the other hand, if some $\II$ is not a subtype of $\II_0$,
\marco{by the newLemma, we know that all its methods are still well typed, and } the generated code in translation has no way to affect the domain of $\II$,\marco{ so rule t-interf can still be applied,} which finishes our proof.\haoyuan{And the well-formedness.}
\end{proof}

### 2015-11-29 paper/sections/Translation.tex
More.
Proof.

\noindent$\begin{array}{l}
\InTextDef{18em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
}{
\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
\methods\ \methods' \cC
}\\
\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \methods'=\ofMethod(\C_0)}
\end{array}$

@@

\subsection{$\ofMethod$}
We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory. To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.

@@

The auxiliary functions are defined below:\\*

>>>>

\noindent$\begin{array}{l}
\InTextDef{18em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
}{
\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
\methods\ \method' \cC
}\\
\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \method'=\ofMethod(\C_0)}
\end{array}$

@@

\subsection{$\ofMethod$}\label{subsec:ofmethod}
We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory. To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.

@@

The auxiliary functions are defined below. Note that we do not need to check if some header is a subtype of what we would generate, this is ensured by $\valid(\C_0)$.
